---
title: "Fruit snack analysis"
author: "Karl Broman"
date: 2015-02-05
output: html_document
---

```{r options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=5,
                      echo=FALSE, results="hide",
                      message=FALSE, warning=FALSE)
barcolor <- "#E6E6FA" # Lavender
```

```{r load_package, include=FALSE}
# load R/broman package (at https://github.com/kbroman/broman)
if(!require(broman)) {
    required <- c("devtools", "RPushbullet", "jsonlite")
    for(pkg in required) {
        if(!require(pkg, character.only=TRUE))
            install.packages(pkg)
    }
    library(devtools)
    install_github("kbroman/broman")
}
```

```{r load_data}
fs <- read.csv("../Data/fruit_snacks.csv")
fs <- fs[,-1] # drop the column with IDs
n_per_package <- rowSums(fs)
```

There were `r nrow(fs)` packages containing a total of `r sum(fs)`
fruit snacks in `r ncol(fs)` colors. (I'd call them "flavors" but we
can't much distinguish among them.) So there was an average of
`r round(sum(fs)/nrow(fs), 1)` fruit snacks per package, with a range of
`r min(n_per_package)` &ndash; `r max(n_per_package)`.  Here's the
distribution:

```{r num_snacks_per_package}
par(mar=c(4.1, 4.1, 1.1, 0.6))
hist(n_per_package, main="", xlab="Number of fruit snacks per package", las=1,
     breaks=seq(min(n_per_package)-0.25, max(n_per_package)+0.25, 0.5),
     col=barcolor)
```

The different colors have quite different frequencies. Here's a plot
of the average number of snacks of each color in a package, with 95%
confidence intervals.

```{r ave_of_each_color, fig.width=6}
par(mar=c(4.1, 4.1, 1.1, 0.6))
ave_per_type <- colMeans(fs)
color <- "slateblue"
ymx <- max(ave_per_type)*1.1
yli <- c(0, ymx*1.05)
x <- seq(along=ave_per_type)
grayplot(x, ave_per_type, type="n",
         xlab="Color", ylab="Ave. pieces per package",
         ylim=yli, hlines=pretty(yli), yaxs="i",
         xlim=range(x)+c(-0.5, 0.5), xaxs="i", xat=NA,
         vlines=x, vlines.lwd=3, vlines.col="gray70",
         bgcolor="gray85")
axis(side=1, at=x, colnames(fs), tick=FALSE, mgp=c(0, 0.2, 0))
xwidth <- 0.2
segments(x-xwidth, ave_per_type, x+xwidth, ave_per_type,
         col=color, lwd=2)
ci <- apply(fs, 2, function(a) t.test(a)$conf.int)
segments(x, ci[1,], x, ci[2,], lwd=2, col=color)
for(i in 1:nrow(ci))
    segments(x-xwidth/2, ci[i,], x+xwidth/2, ci[i,], lwd=2, col=color)
```

Here are histograms of the number of snacks per package for each color.

```{r hist_of_each_color, fig.height=12}
par(mfrow=c(5,1), mar=c(4.1, 4.1, 2.1, 0.6))
br <- seq(-0.25, max(fs)+0.25, by=0.5)
for(i in 1:ncol(fs))
  hist(fs[,i], breaks=br, main=colnames(fs)[i], xlab="Number per package",
       las=1, col=barcolor)
```

We can see immediately that there is variation in the frequency of the
different colors.  But the distributions look plausibly random. If
colors were randomly assigned to packages (but at color-specific
frequencies), these would be mixtures of binomial
distributions. (_Mixtures_, because the number of snacks in a package
varies.)

I'll first focus on the purple ones.

[**More to come.**]
